---
title: "DSCI353-353m-453: LE2b TSA"
subtitle: "2001-353-353m-453-LE2b"
author: "Roger H. French, Peitian Wang"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    latex_engine: xelatex
    toc: true
    number_sections: true
    toc_depth: 6
    highlight: tango
  html_document:
    toc: yes
urlcolor: blue
---

<!--
# Script Name: 
# Authors: Roger H. French
# License: Creative Commons Attribution-ShareAlike 4.0 International License.
##########
# Latest Changelog Entires:
# v0.00.01 - 1501cwru-dsci-NAMEIT.RMD - Roger French started this blank Rmd
##########

# Rmd code goes below the comment marker!
-->


\setcounter{section}{2}
\setcounter{subsection}{2}
\setcounter{subsubsection}{0}

#### Lab Exercise 2b: Time Series Analysis

2 Questions, for a total of 3 points

  - Question 1 = 1 points
  - Question 2 = 2 points

Time series are a common type of data, 

  - consisting of measurements that are continuous over a time range. 

In this project we will be using classical decomposition 

  - to perform analysis on a time series.

First as an introduciton to decomposition we will have a quick example.

##### Question 1

What is the decomposition of a timeseries?

  - The AirPassangers data set of airline passangers every month for 12 years

```{r}
library(datasets)
library(dplyr)
library(ggplot2)
air <- as.data.frame(AirPassengers)

airpass <- AirPassengers
airp.dcpM <- decompose(airpass, type = "multiplicative")
plot(airp.dcpM)
```

Plot the total time series of air passangers 

  - What do you notice about the plot?

```{r}

plot.ts(air)
"There is a repetitive or close enough occurance of number of passenger for several month.
Also, there is a local maximum occuring every several month with a overall trending of increasing."
```

Use the ts() function in base R 

  - to define AirPassangers as a time series with a yearly trend 

If the data is taken monthly, 

  - what will the frequency (points per season) of a yearly season be? 

```{r}

?ts()
airts <- ts(air, frequency = 12)

```

Use the decompose() function 

  -to demcompose the time series and remove the seasonality

The type for this time series is multiplicative

  - Plot the decomposed time series, what do you notice about the trend?

The trend of number of air passangers was keep increasing.

```{r}

airpass <- AirPassengers
airp <- decompose(airpass, type = "multiplicative")
plot(airp)

```

Isolate the trend and plot the trend on top of the raw data with the seasonality included

  - How well does the trend represent the data?

```{r}

ts_air <- ts(air, frequency = 12)
dcp_air = decompose(ts_air, "multiplicative")

plot(dcp_air$trend)
plot(ts_air)
plot(dcp_air$seasonal)



```

##### Question 2

Now lets try this with a real world time series. 

We'll be using one month of power and weather data from a solar power plant.

The data set variables are as follows:

  - time: The timestamp at which the data was taken
  - ghir: Global Horizontal Irradiance from a sensor at the site, 
    - the power from the sunlight over an area normal to the surface of the earth $(Watts/m^2)$
  - iacp: The AC power from the power plant $(kW)$
  - temp: The air temperature $(Celsius)$
  - ghi_solargis: The Global Horizontal Irradiance, not from a sensor, 
    - but predicted using weather modeling $(Watts/m^2)$
  - clear: A logical indicating whether the sky was "clear" during measurement, 
    - determined by comparing the ghi and ghi_solargis data
  - ratio: the ratio of the Global Horizontal Irradiance 
    - and the Plane of Array Irradiance (the irradiance normal to the surface of a tilted module)

The power from solar panels is dependant on the irradiance hitting it, 

  - so a power reading is often meaningless without a corresponding irradiance measurement.

It is useful to have multiple sources of irradiance measurements. 

Sensors on the ground are useful because 

  - they strongly represent the irradiance that is hitting the module; 
  - however, sensors can begin to drift if not cleaned or calibrated properly. 
  - An unstable sensor can render an entire data set useless.

To combat this, we also have irradiance data from SolarGIS, 

  - a company that uses satellite images to model and predict 
  - the irradiance at the surface of the earth. 

  - Plot the irradiance and power for the first week of data, 
    - how does the irradiance look compared to 
      - what you would expect from the trend of sunlight? 
    - How well does the power and irradiance match up?

```{r}
pnw <- read.csv("C:/Users/k0063/cwrudsci-20s-dsci353-353m-453-prof-f678d9ea895a/1-assignments/lab-exercise/LE2/data/2001-353-353m-453-le2b-tsa-data-final.csv", skip = 1, sep = ",", header = TRUE)
 
irrad <- select(pnw, ghir)
ts_ir <- ts(irrad, start = 1, end = 672)
plot(ts_irradiance)
power <- select(pnw, iacp)
ts_power <- ts(power, start = 1, end = 672)
plot(ts_power)

```

Use the ts() functions and the stlplus() function from the stlplus package 

  - to decompse the sensor and SolarGIS irradiance and the power 
    - for the whole month. 
  -  Plot each of the decompositions, what do you notice?

```{r}
# think carefully about the frequency you'll need to define for this data
# what is the seasonal component to this data and how nay data points make up a season?
# use s.window = "periodic" for the stlplus function
library(stlplus)
?stlplus()

sensor <- ts(irradiance, frequency = 2, start = 1, end = 2880)
sen_stl <- stl(sensor, s.window = "periodic")
plot(sen_stl)

Solgis <- select(powerandweather, ghi_solargis)
solg <- ts(Solgis, frequency = 2, start = 1, end = 2880)
Sol_stl <- stl(solg, s.window = "periodic")
plot(Sol_stl)

pow <- ts(power, frequency = 2, start = 1, end = 2880)
pow_stl <- stl(pow, s.window = "periodic")
plot(pow_stl)



```

Isolate the trends for the 3 time series you just decomposed 

  - and build a linear model for each one.

Compare the models to each other, 

  - how are they different?

```{r}

plot(sensor_stl$trend)

```

Solar panel degradation leads to less power output over time 

  - at the same irradiance conditions.

Based on the linear models you found for the trends of power and irradiance, 

  - is this system degrading over time?

How do the sensor GHI and the SolarGIS GHI compare to power?





#### Links
 
http://www.r-project.org 

http://rmarkdown.rstudio.com/  

<!--
# Keep a complete change log history at bottom of file.
# Complete Change Log History
# v0.00.00 - 1405-07 - Nick Wheeler made the blank script
##########

