---
title: "DSCI353-353m-453: LE1b  "
subtitle: "2001-353-353m-453-LE1b-TidyText-NAME.Rmd"
author: "Roger H. French, Peitian Wang"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    latex_engine: xelatex
    toc: true
    number_sections: true
    toc_depth: 6
  html_document:
    font-size: 10em
    self_contained: yes
    toc: yes
urlcolor: blue
---
 
<!--
# Script Name: 1906-tu-itfun-adswr-LE1b-TidyText-NAME
# Purpose: This is a blank Rmd file to start a new open science data analysis from
# Authors: Roger H. French
# License: Creative Commons Attribution-ShareAlike 4.0 International License.
##########
# Latest Changelog Entires:
# v0.00.01 - 1501cwru-dsci-NAMEIT.RMD - Roger French started this blank Rmd
##########

# Rmd code goes below the comment marker!
-->


\setcounter{section}{1}
\setcounter{subsection}{1}
\setcounter{subsubsection}{0}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

```

### LE1b, 3 points, 6 questions.

  -  Each Question 0.5 points

#### Details

  - Due Wednesday June 12th
    - Before Class
  - The grading is done on how you show your thinking, 
    - explain yourself and 
    - show your Rcode and 
    - the output you got from your code.
  - Code style is important
    - Follow Rstudio code diagnostics notices
    - And the [Google R Style Guide](https://google.github.io/styleguide/Rguide.xml)

#### Text Mining of Song Lyrics

  - Complete the given problems
  - dplyr, ggplot, and pipelines are highly recommended
  - We will be using the Tidytext package to aid with our text mining
  - The dataset for this assignment is a collection of the information and lyrics from every top 100 billboard song since 1965

### Problem 1

  - Load in the data set
  - Print the lyrics of the #4 song from 1988
  - Show the top 20 artists with the most hits, mo

```{r}
  ##ANS:

billboard<- read.csv('/Users/k006360/Documents/CWRU/R/cwrudsci-20s-dsci353-353m-453-prof-acfeb82427e8/1-assignments/lab-exercise/LE1/data/billboard_lyrics_1964-2015.csv')
  library(dplyr)
  billboard %>% filter(Year=='1988') %>% filter(Rank=='4')
  
artist <- billboard %>% arrange(Artist) %>% count(Artist)
  top20th <- artist %>% arrange(n)
  top20th <- artist %>% filter(n>12)
  top1st <- top20th %>% filter(n==max(n))
```

### Problem 2
p
  - Build a histogram of the amount of times artists appear on the top hits billboard, what does the trend look like, what does it suggest about 1 hit wonders?

```{r}
  ##ANS:
library(ggplot2)
billboard %>% ggplot(sort(Artist,decreasing = FALSE), mapping=aes(Artist)) + geom_histogram(stat='count',aes(fill=factor(Year)))

monohitwonder <- artist %>% filter(n==1)

```

### Problem 3

  - Lets do a lyrical comparison between 2 top artists, Elton John and Eminem
  - First, filter the main data set to only the 2 given artists (you can make them seperate data frames or keep them together, your call)
  - Use the unnest_tokens() function from tidytext to split the lyrics up so that each word has its own row
  - Show the top 10 most commonly used words for each artist

```{r}
library(dplyr)
L <- billboard %>% filter(Artist=="elton john")
M <- billboard %>% filter(Artist=='eminem')
Llyric <- L %>% select(Lyrics)
Mlyric <- M %>% select(Lyrics)
  Ltxt <- Llyric %>% tibble(txt='Lyrics')
  Mtxt <- Mlyric %>% tibble(txt='Lyrics')
  l<-Ltxt %>% unnest_tokens(word, txt)
  m<-Mtxt %>% unnest_tokens(word,txt)

```

  - It should be clear that most of these are not significantly meaningful words and should be removed, we can remove them using the stop_words dataframe provided with the tidytext package (hint: look at dplyr anti_join())
  - Remove them and show the top ten remaining words for each artist

```{r}

data("stop_words")



```

### Problem 4

  - Build a word cloud of the lyrics for each atrist, find an R package that will help you with this
  - Compare and contrast the word clouds
  - Did the stop_words dataframe work well to remove non-meaningful words?

```{r}

library(wordcloud)



```

### Problem 5

  - The tidytext package gives us the ability to run sentiment analysis as well
  - There are multiple sentiment methods available but let's just use positive and negative sentiment

```{r}

bing <- get_sentiments("bing")

```

  - What is the bing dataframe?
  - Pull out the "positive" and "negative" words for each artist (hint: dplyr inner_join())
  - For each song per artist, determine the next positive or negative sentiment, there are several ways to do this - the most straightforward being assign a 1 to a positive lyric and -1 to a negtive lyric then sum them together for each song
  - Make a bar plot of the net sentiment of each song for each artist, make these plots high quality as well
    - Properly name the axes and title each plot
    - Color by whether the song is overall positive or negative
    - rotate the x-axis names so long song titles are legible
    - Arrange the bars in accending or descending order

```{r}


```
 
  - What differences do you notice between the 2?
  - Based on what you might know about some of these songs, do you think the sentiment analysis gives a good indication of the positive of negative tone for each song or not?
  - If you're interested, play around with this data set, see if your favorite artist is in here and see if you can find anything else


### Problem 6

  - Now that you are familiar with this data set, since you've been exploring it
  - Define your own question, and do some analysis to answer that question
  - Look through the R packages you have already installed to get some ideas
  - And for the answer, provide a good visualization

#### Question: 

```{r}

library(wordcloud)



```

#### Answer: 