---
title: "2001-LE3b-ClusteringLifeExpectancy"
subtitle: "2001-353-353m-453-LE3b-ClusteringLifeExpectancy"
author: "Roger H. French, Peitian Wang"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    latex_engine: xelatex
    toc: TRUE
    number_sections: TRUE
    toc_depth: 6
    highlight: tango
  html_document:
    toc: TRUE
urlcolor: blue
---

1 Questions, 1 points total.

#  the best number of clusters in the life.expectancy.1971 dataset

We will use the life.expectancy.1971 dataset 

  - about life expectancy in several countries in 1971, 
  - which is part of the cluster.datasets package. 

It includes 10 attributes: 

  + the country were the data has been collected, 
  + the year of data collection, and 
  + the life expectancy (remaining) for 
    - male and female individuals 
    - aged 0 years old, 25, 50, and 75. 
  
As with the previous crime dataset, 

  - this one also does not specify the membership of our cases to categories. 

So again, we will have to decide on the number of clusters by ourselves. 

  + e will examine how to do so more precisely. 
  + We will create a function for this purpose. 

Before we do that, let's discover the dataset we will use. 

Let's start by loading and examining to dataset.

Show a summary of the dataset

```{r}
library(cluster.datasets)
data(life.expectancy.1971)
summary(life.expectancy.1971)
```

## Cleaning the data: Madagascar, Cameroon, Trinidad, US

Even without computing the mean and standard deviations for the variables, 

  - we can notice that there is quite some variation regarding life expectancy 
  - (please refer to the complete output on your screen as well). 

A first observation, which is broadly documented, 

  - is that women have a longer remaining life expectancy 
  - than men, at all ages. 

A country strikes in this list—in Madagascar, 

  - at the time of data collection, 
    - women apparently did not have longer life expectancy 
    - than men in their young and old years. 

Further, the mean life expectancy at birth 

  - was only 38 
  - for both women and men. 

This is also the life expectancy of females in Cameroon at that time, 

  - whereas males were expected to live even a little less (34 years). 

Looking at the table, 

  - we can notice that Trinidad and the US are entered several times, 
  - as data collection was carried out more than once. 

We will therefore discard case 23 (the second entry for Trinidad), 

  - as well as both cases 24 and 27 (US, data collected in 1966 and 1967) 
  - because cases 25 and 26 are more specific, 
  - as they provide estimations for White and Nonwhite individuals. 

Let's create a new dataset without these cases 

  - before we proceed with cluster analysis.

Best done using dplyr

```{r}
life <- life.expectancy.1971
life <- life[-c(23,24,27),]

```

## Scale the Data, and Add Some Attributes

Here we will scale the data. 

The importance of scaling data 

  - has been discussed in the first section of this chapter. 

Question: Why do we need to scale this data?

  - Answer: To find the pattern of the data much easier.

We also add some attributes to the dataset, 

  + corresponding to the ratio of male life expectancy 
    - to female life expectancy at all ages, 
  + as the difference between male and females 
    - would be lost in data scaling (all means will be 0).

Use cbind for this.

```{r}
life <- life %>% mutate(r0 = m0 / f0) %>% mutate(r25 = m25 / f25) %>% mutate(r50 = as.numeric(m50) / as.numeric(f50)) %>% mutate(r75 = as.numeric(m75) / as.numeric(f75))

```

When you run your code, you will notice an error. 

It happens that attribute f50 

  - is composed of strings 
    - instead of numeric values 
  - (type mode(life$f50) to check this). 

This is a type of problem you might encounter 

  - when dealing with data you have not prepared yourself 
    - (and sometimes even with your data). 

The solution is obviously 

  - to convert the attribute to numeric values 
  - before being able to compute the ratios.

Use as.numeric, then cbind

```{r}
mode(life$f50)

life <- life %>% mutate(r0 = m0 / f0) %>% mutate(r25 = m25 / f25) %>% mutate(r50 = as.numeric(m50) / as.numeric(f50)) %>% mutate(r75 = as.numeric(m75) / as.numeric(f75))


```

We can now repeat our assignment to life.temp with a successful result, 

  - and scale the data frame (omitting rows 1 and 2: 
    - name of country and year of data collection). 

We first convert to a data frame 

  - to get rid of information about mean and standard deviation 
  - that is contained in the returned object; we then convert to a matrix again.

```{r}
lscale <- life[,-c(1,2)]
lscale_ <- as.matrix(data.frame(lscale))
```

## Now lets do some External Validation of our Classification

When examining the iris dataset, 

  - we had the correct solution 
    - regarding the number of clusters 
    - and the classification of cases. 

This is not the case here

  - we can not tell before running the analyses 
    - the number of groups in our data. 

We will therefore rely on computational trickery to discover them; 

  - cluster analysis will be performed iteratively 
    - and the clustering solutions will be compared 
    - using several indexes for determining the ideal number of clusters.
    
More information about such indexes can be found in the paper 

  - [Experiments for the number of clusters in k-means, by Chiang and Mirkin, 2007](https://link.springer.com/chapter/10.1007%2F978-3-540-77002-2_33) 
  - (it should also be in your readings). 

Here we rely on NbClust() function 

  - from the NbClust package, 
  - which we install and load:

```{r}
library(NbClust)
mode(lscale)
```

We simply call the NbClust() function 

  - specifying the data and clustering algorithm to be used.
  
By default, the function will perform clustering 

  - using the Euclidean distance 
  - and compute all available indexes. 
  
The reader is advised to consult the package help documentation

  - for more information about customization.

```{r}
lnum <- unlist(lapply(lscale, as.numeric))
rgxn <- NbClust(lnum, method = "kmeans",distance = "euclidean" ,index = "all")
```

Question: This shows that how many clusters is the most appropriate solution.

  -  Answer: the best number of clusters is  5
  
Question: Explain how/what shows this, in the NbClust output.

  - Answer: The Hubert index is a graphical method of determining the number of clusters.In the plot of Hubert index, we seek a significant knee that corresponds to a significant increase of the value of the measure i.e the significant peak in Hubert index second differences plot. 
 
*** : The D index is a graphical method of determining the number of clusters. In the plot of D index, we seek a significant knee (the significant peak in Dindex second differences plot) that corresponds to a significant increase of the value of the measure. 




# Cites

* Gareth James, Daniela Witten, Trevor Hastie, and Robert Tibshirani. An Introduction to Statistical Learning: With Applications in R. 1st ed. 2013, Corr. 5th printing 2015 edition. Springer Texts in Statistics. New York: Springer, 2013.

