---
title: "DSCI353-353m-453: LE2a Classification"
subtitle: "2001-353-353m-453-LE2a"
author: "Roger H. French, Peitian Wang"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    latex_engine: xelatex
    toc: true
    number_sections: true
    toc_depth: 6
    highlight: tango
  html_document:
    toc: yes
urlcolor: blue
---

<!--
# Script Name: 
# Authors: Roger H. French
# License: Creative Commons Attribution-ShareAlike 4.0 International License.
##########
# Latest Changelog Entires:
# v0.00.01 - 1501cwru-dsci-NAMEIT.RMD - Roger French started this blank Rmd
##########

# Rmd code goes below the comment marker!
-->


\setcounter{section}{2}
\setcounter{subsection}{1}
\setcounter{subsubsection}{0}

#### Lab Exercise 2a: Classifiation

2 Questions, for a total of 2 points

  - Question 1 = 1.5 points
  - Question 2 = 0.5 points

#### Question ISLR Chapter 4, Exercise 10

This question should be answered using the Weekly data set, 

  - which is part of the ISLR R package. 
  
This data is similar in nature to the Smarket data 

  - from this chapter’s lab, 
  - except that it contains 
    - 1,089 weekly returns 
    - for 21 years, 
    - from the beginning of 1990 to the end of 2010.
    
##### (a) Produce some numerical and graphical summaries of the Weekly
data. 

Do there appear to be any patterns?

```{r}
 ggplot(data=Weekly,aes(x=Volume,y=Lag1))+geom_line(aes(group = factor(Year), color= factor(Year)))+geom_point(aes(color=factor(Year)))
ggplot(data=Weekly,aes(x=Volume,y=Lag2))+geom_line(aes(group = factor(Year), color= factor(Year)))+geom_point(aes(color=factor(Year)))
ggplot(data=Weekly,aes(x=Volume,y=Lag3))+geom_line(aes(group = factor(Year), color= factor(Year)))+geom_point(aes(color=factor(Year)))
ggplot(data=Weekly,aes(x=Volume,y=Lag4))+geom_line(aes(group = factor(Year), color= factor(Year)))+geom_point(aes(color=factor(Year)))
ggplot(data=Weekly,aes(x=Volume,y=Lag5))+geom_line(aes(group = factor(Year), color= factor(Year)))+geom_point(aes(color=factor(Year)))

W1 <- Weekly %>% mutate(Lag = Lag1+Lag2+Lag3+Lag4+Lag5)
ggplot(data=W1,aes(x=Volume,y=Lag))+geom_line(aes(group = factor(Year), color= factor(Year)))+geom_point(aes(color=factor(Year)))
ggplot(data=W1,aes(x=Volume,y=Lag))+geom_line(aes(group = factor(Year), color= factor(Year)))+geom_point(aes(color=factor(Year)))
ggplot(data=W1,aes(x=Volume,y=Lag))+geom_line(aes(group = factor(Year), color= factor(Year)))+geom_point(aes(color=factor(Year)))


summary(Weekly)
plot(Weekly)
cor(Weekly[,-9])
table(Weekly$Direction)/sum(table(Weekly$Direction))
head(Weekly[,c(8,9)])

```

###Answer: 
            Yes, the Volume increase while the Year increase, and the deviation of Lag (including Lag1~Lag5 and the sum of Lag1 ~ Lag5) is more divergent while the Year increase.


##### (b) Use the full data set to perform a logistic regression 

  - with Direction as the response 
  - and the five lag variables plus Volume as predictors. 

Use the summary function to print the results. 

Do any of the predictors appear to be statistically significant? 

If so, which ones?

```{r}
glm.fit <- glm(Direction~.-Year-Today,data=Weekly,family="binomial")
summary(glm.fit)


```

###Answer:  The only statistically significant variables are Lag1, and Lag2.


##### (c) Compute 

  - the confusion matrix and 
  - overall fraction of correct predictions. 

Explain what the confusion matrix is telling you

  - about the types of mistakes made by logistic regression.
  
```{r}
glm.probs <- predict(glm.fit,type = "response")
glm.pred <- rep("Down",nrow(Weekly))
glm.pred[glm.probs>0.5] = "Up"

table(glm.pred,Weekly$Direction)

mean(glm.pred == Weekly$Direction)
print(558/(558+47))
print(56/(428+56))

```

###Answer: The algorithm classifies about 56% observations are correct. It classified on the days that the stock increases about 92% of ‘Up’ entries are correct. It classified poorly on days the market is 'Down', with only 11% are correct.


##### (d) Now fit the logistic regression model 

  - using a training data period from 1990 to 2008, 
  - with Lag2 as the only predictor. 
  
Compute 

  - the confusion matrix and 
  - the overall fraction of correct predictions
  - for the held out (testing) data 
    - (that is, the data from 2009 and 2010).
    
###Answer:    
```{r}
train <- Weekly[,"Year"] <= 2008

glm.fit <- glm(Direction~Lag2,data = Weekly,subset = train, family = "binomial")
summary(glm.fit)

glm.probs <- predict(glm.fit,Weekly[!train,],type = "response")

glm.pred <- rep("Down",nrow(Weekly))
glm.pred[glm.probs>0.5] = "Up"

table(glm.pred,Weekly[,"Direction"]) 

mean(glm.pred == Weekly[,"Direction"])

```



##### (e) Repeat (d) using LDA.
###Answer: 
```{r}
library(MASS)

lda.fit <- lda(Direction~Lag2,data=Weekly,subset=train)
lda.fit

lda.pred <- predict(lda.fit,Weekly[!train,])
lda.class <- lda.pred$class

table(lda.class,Weekly[!train,9])

mean(lda.class == Weekly[!train,9])
```




##### (f) Repeat (d) using QDA.

###Answer: 
```{r}
qda.fit <- qda(Direction~Lag2,data=Weekly,subset=train)
qda.fit

qda.pred <- predict(qda.fit,Weekly[!train,])
qda.class <- qda.pred$class

table(qda.class,Weekly[!train,9])

mean(qda.class == Weekly[!train,9])
```





##### (g) Repeat (d) using KNN with K = 1.

###Answer: 
```{r}
library(class)

train.X <- cbind(Weekly[train,3])
test.X <- cbind(Weekly[!train,3])

train.Direction <- Weekly[train,c(9)]
test.Direction <- Weekly[!train,c(9)]

knn.pred <- knn(train.X,test.X,train.Direction,k=1)

table(knn.pred,test.Direction)

mean(knn.pred == test.Direction)
```




##### (h) Which of these methods 

  - appears to provide the best results on this data?
  
  
(i) Experiment with different combinations of predictors, 

  - including possible transformations 
  - and interactions, 
  - for each of the methods. 
  
Report the variables, method, and associated confusion matrix 

  - that appears to provide the best results on the held out data. 
  
Note that you should also experiment 
  
  - with values for K in the KNN classifier.

###Answer: 
```{r}
glm.fit <- glm(Direction~Lag2,data = Weekly,subset = train, family = "binomial")
summary(glm.fit)

glm.probs <- predict(glm.fit,Weekly[!train,],type = "response")

glm.pred <- rep("Down",nrow(Weekly))
glm.pred[glm.probs>0.55] = "Up"

table(glm.pred,Weekly[,"Direction"])

mean(glm.pred == Weekly[,"Direction"])

glm.fit <- glm(Direction~Lag2+I(Lag2^2)+I(Lag2^3),data = Weekly,subset = train, family = "binomial")
summary(glm.fit)

glm.probs <- predict(glm.fit,Weekly[!train,],type = "response")

glm.pred <- rep("Down",nrow(Weekly))
glm.pred[glm.probs>0.5] = "Up"

table(glm.pred,Weekly[,"Direction"])

mean(glm.pred == Weekly[,"Direction"]) 

glm.fit <- glm(Direction~sqrt(abs(Lag2)),data = Weekly,subset = train, family = "binomial")
summary(glm.fit)

glm.probs <- predict(glm.fit,Weekly[!train,],type = "response")

glm.pred <- rep("Down",nrow(Weekly))
glm.pred[glm.probs>0.5] = "Up"

table(glm.pred,Weekly[,"Direction"])

mean(glm.pred == Weekly[,"Direction"])

glm.fit <- glm(Direction~log10(abs(Lag2)),data = Weekly,subset = train, family = "binomial")
summary(glm.fit)

glm.probs <- predict(glm.fit,Weekly[!train,],type = "response")

glm.pred <- rep("Down",nrow(Weekly))
glm.pred[glm.probs>0.5] = "Up"

table(glm.pred,Weekly[,"Direction"])

mean(glm.pred == Weekly[,"Direction"])

train.X <- cbind(Weekly[train,3])
test.X <- cbind(Weekly[!train,3])

train.Direction <- Weekly[train,c(9)]
test.Direction <- Weekly[!train,c(9)]


errors <- c()

maxK <- 100
step <- 2

for(j in seq(1,maxK,step)){
  knn.pred <- knn(train.X,test.X,train.Direction,k=j)
  table(knn.pred,test.Direction)
  errors <- c(1-mean(knn.pred == test.Direction),errors)
}

data <- cbind(seq(1,maxK,step),errors)
plot(data,type="l",xlab="k")

knn.pred <- knn(train.X,test.X,train.Direction,k=which.min(errors))
table(knn.pred,test.Direction)

mean(knn.pred == test.Direction)
```




#### Question ISLR Chapter 4, Exercise 12


This problem involves writing functions.

##### (a) Write a function, Power(), 

  - that prints out the result of raising 2 to the 3rd power. 
  
In other words, your function should 

  - compute $2^3$ and print out the results.
  
  
Hint: Recall that x^a raises x to the power a. 

  Use the print() function to output the result.

###Answer:  
```{r}
Power = function(){2^3}
Power()
```

   

(b) Create a new function, Power2(), 

that allows you to pass any two numbers, 

  - x and a, 
  - and prints out the value of x^a. 

You can do this by beginning your function with the line 

  - `Power2 = function (x , a ) {`
  
You should be able to call your function by entering, 

  - for instance,
  - `Power2 (3 ,8)` on the command line. 
  
This should output the value of $3^8$, 

  - namely, 6,561.

###Answer: 
```{r}
Power2 = function(x,a){x^a}
Power2(3,8)
```


  
##### (c) Using the Power2() function that you just wrote, 

  - compute $10^3$, $8^17$, and $131^3$.

###Answer:   
```{r}
Power2(10,3)
Power2(8,17)
Power2(131,3)
```




(d) Now create a new function, Power3(), 

  - that actually returns the result $x^a$ as an R object, 
  - rather than simply printing it to the screen. 
  
That is, if you store the value $x^a$ 

  - in an object called result within your function, 
  - then you can simply return() this result, 
  - using the following line:
    - `return ( result )`
    
The line above should be 

  - the last line in your function, 
  - before the } symbol.

###Answer: 
```{r}
Power3 = function(x,a){ result=x^a; return(result)}
```



##### (e) Now using the Power3() function, 

  - create a plot of $f(x) = x^2$.
  - use ggplot2
  
The x-axis should display 

  - a range of integers from 1 to 10, 
  
and the y-axis should 

  - display $x^2$. 
  
Label the axes appropriately,

  - and use an appropriate title for the figure. 
  
Consider displaying either 

  - the x-axis, the y-axis, 
  - or both on the log-scale. 

You can do this by using 

  - log=‘‘x’’, log=‘‘y’’, or log=‘‘xy’’ 
    - as arguments to the plot() function.
  - but do it using ggplot2.

###Answer: 
```{r}
x = 1:10
y = sapply(x,function(x){Power3(x,3)})

plot(x,y,log='x',main='x ~ x^2',type='l')
```

  
  
##### (f) Create a function, PlotPower(), 

  - that allows you to create a plot 
  - of $x$ against $x^a$ 
  - for a fixed a 
  - and for a range of values of x. 
  
For instance, if you call

  - `PlotPower (1:10 ,3)`
  - then a plot should be created with 
    - an x-axis taking on values 1, 2, . . . , 10, 
    - and a y-axis taking on values 13 , 23 , . . . , 103 .

###Answer: 
```{r}
PlotPower = function(x,a){
  y = sapply(x,function(x){x^a})
  plot(x,y,main=paste('x~x^',a,sep = ''),type='l')
}
PlotPower(1:10,3)
```




#### Links
 


<!--
# Keep a complete change log history at bottom of file.
# Complete Change Log History
# v0.00.00 - 1405-07 - Nick Wheeler made the blank script
##########

